\Large \bf Cyclone: High Availability for Persistent Key Value Stores

### 2017-08-08 paper/paper.tex
Updates.

@@

Cyclone uses the RAFT protocol~\cite{raft} to keep the logs in sync across the
replicas. On
failure of the leader replica, a new leader is automatically elected -- RAFT
ensures that the new leader has the most up to date log. The Cyclone client
library automatically locates the new leader on a failover, resulting in only a
brief interruption in service for clients. Although read requests (GET
operations) are not logged, they are serviced in FIFO order with the write
requests (PUT/DELETE operations) in Cyclone's log at the moment the
read request is received. This ensures linearizability in the face of failovers.

@@

The two level log design in Cyclone lays the foundation for improvements to the
network stack that we discuss next. We use the RAFT protocol to replicate the
NVM log from
the leader replica to the followers. Each replica \emph{independently} drains
committed (in terms of RAFT consensus) entries from the NVM log to the
flashlog. Block IO therefore is no longer a consideration when optimizing data
movement over the network, unlike systems that need to deal with such
problems~\cite{reflex}. The task of replication reduces to that of efficiently
sending a block of data already present in directly attached memory over the
network.

\section{Network}
\label{sec:network}

@@

There is an in progress transformation of durable memory in datacenters
with the advents of 3DXPoint. Researchers have speculated on the impact of cheap
and high volume directly attached 3DXPoint~\cite{tiering}. Memory vendors have
also begun shipping software libraries such as NVML~\cite{nvml} for building durable
in-memory data structures using undo logging. A durable key value store using
3DXPoint would be an idea target for Cyclone because each NVM log entry can be
discarded as soon as it is applied to the persistent key value store removing
the need for the flashlog. This would allow Cyclone to operate at pure packet
switching speeds, a goal we hope to achieve in the future with a single level
log for NVML data store applications.

@@

\section{Conclusion}
Cyclone addresses two fundamental difficulties in adding high availability to
persistent key value stores. It leverage a small amount of non-volatile memory
to avoid paying the cost of the block IO interface for every update
operation. It also leverages software packet switching principles to replicate
the write ahead log, while making the best possible use of commodity networking
hardware. These techniques enable Cyclone to provide high availability to
persistent key value stores, without compromising on their performance. 
\newcommand\myurl[2]{\url{#1}}
\bibliographystyle{plain}
\bibliography{paper}

>>>>

These two performance bottlenecks illustrated in Figure~\ref{fig:problem} are
not \emph{independent}. In particular, durability semantics mean that the
key-value store  cannot respond to the client until the operation has been
persisted to the log, since it must be completed even on a crash and subsequent
recovery. Therefore, addressing the network bottleneck is pointless without also
addressing the storage bottleneck. Cyclone shows how a small amount of
non-volatile memory can be used to address both bottlenecks.

@@

Cyclone uses the RAFT consensus protocol~\cite{raft} to keep the logs in sync
across the replicas. On failure of the leader replica, a new leader is
automatically elected -- RAFT ensures that the new leader has the most up to
date log. The Cyclone client library automatically locates the new leader on a
failover, resulting in only a brief interruption in service for
clients. Although read requests (GET operations) are not logged, they are
serviced in FIFO order with the write requests (PUT/DELETE operations) in
Cyclone's log at the moment the read request is received. This ensures
linearizability in the face of failovers.

@@

In this section we demonstrated how Cyclone uses a small amount of non-volatile
memory to improve storage bottlenecks when maintaining a durable log on a single
machine. In the next section, we deal with replicating this log efficiently
across multiple machines.

\section{Network}
\label{sec:network}
Cyclone uses the RAFT~\cite{raft} consensus protocol to replicate the log of
operations across machines. It promises the client that any operation responded
to is durable and its results \emph{available} as long as a majority of replicas
continue to function. Cyclone's failure model therefore is \emph{fail-recovery},
where nodes that can recover from failure simply continue the protocol. This
requires RAFT to persist every log entry \emph{before} it sends it out for
replication. Storage overheads are therefore not independent of network
overheads and we must also deal with the storage overhead to improve the
performance of replication.

We leverage the two level log from the previous section to remove storage
bottlenecks from the critical path of replication in much the same way we did
for logging on a single machine. RAFT is only aware of and replicates the top
level NVM log. Each replica \emph{independently} drains committed (in terms of
RAFT consensus) entries from the NVM log to the flashlog. Block IO therefore is
no longer a consideration when optimizing data movement over the network, unlike
systems that need to deal with such problems~\cite{reflex}. The task of
replication reduces to that of efficiently sending a block of data already
present in directly attached memory over the network.

@@

Database practitioners have considered the utility of byte adressable
non-volatile memory in improving logging for
databases~\cite{nvram_log}. However, that work proposes placing the entire log
in NVM, presuming sufficient availability of such memory. It also does not
consider replication.

@@

\section{Conclusion}
Cyclone shows how one can leverage a small amount of non-volatile memory to
address two fundamental difficulties in adding high availability to persistent
key value stores. It leverage a small amount of non-volatile memory to avoid
paying the cost of the block IO interface for every update operation. This
removes block IO latency from the task of appending updates to the log as well
as sending it out over the network for replication. In turn, this enables us to
apply software packet switching techniques and concurrency to improve the
performance of and make full use of available resources when replicating log
entries over the network. This enables Cyclone to provide high availability to
persistent key value stores, without compromising on their performance. 
\newcommand\myurl[2]{\url{#1}}
\bibliographystyle{plain}
\bibliography{paper}
